Trauma treatment delay

### 2023-12-04 manuscriptlongversion.Rmd
text

The outcome is delayed treatment, as identified by the multidisciplinary
review board and recorded in the trauma care quality database. The
outcome we'll be using is "long time untill CT" and "long time until
operation", which will be referred to as delayed treatment from now on.

@@

The outcome is delayed treatment as identified by the multidisciplinary
review board and recorded in the trauma care quality database. The
independent variables are gender, GCS on arrival, respiratory rate on
arrival, systolic blood pressure on arrival, number of injuries, ISS and
age. Of these variables, GCS, respiratory rate, systolic blood pressure,
number of injuries, ISS and age are used as continuous variables in the
analysis. Gender is used as a categorical variable.

@@

Rstudio was used to perform the logistical regression. The code was
written to clean and prepare the data from the database. The data was
processed to have the variables customized to fit into the logistical
regression model, such as making them into factors, renaming outcomes in
the database and removing unknown values. The variables were either
binary or continuous. This code was used on scrambled data at first so
to avoid selection bias as much as possible. When the code was finished,
it was run on the unscrambled data.

@@

For the missing hospital values in GCS, RR and SBP, values from EMT
personnel were used. This was accomplished by identifying missing data
in the column for the hospital value of the variable and, if available,
replacing the row in the dataset with the prehospital value. If a
prehospital value is missing, the patient is excluded from the study.

@@

As we were running the logistical regression, further were excluded from
the study due to missing data. Among the 8,312 patients screened for
ofi, 7,378 were included in the study. 934 patients were excluded from
the study due to missing values in one or more variables.

@@

The biases that might have been present in this study would be selection
bias and misclassification bias.

@@

21

>>>>

The outcome is the ofi delayed treatment. An ofi is any failure in care,
e.g. clinical judgement error, any preventable or potentially
preventable death, missed diagnosis, technical error and delayed
treatment as identified and decided my the MoM conference. The outcomes
we'll be using is "long time until CT" and "long time until operation",
which will be referred to as delayed treatment from now on.

@@

The independent variables are gender, GCS on arrival, respiratory rate
on arrival, systolic blood pressure on arrival, number of injuries, ISS
and age. Of these variables, GCS, respiratory rate, systolic blood
pressure, number of injuries, ISS and age are used as continuous
variables in the analysis. Gender is used as a categorical variable.

For missing hospital values in GCS, RR and SBP, values from EMT
personnel were used. This was accomplished by identifying missing data
in the column for the hospital value of the variable and, if available,
replacing the row in the dataset with the prehospital value. If a
prehospital value is missing, the patient is excluded from the study.

@@

Rstudio was used to perform the logistical regression. The code was
written to clean and prepare the data from the database. The data was
processed to have the variables customized to fit into the logistical
regression model, such as making them into factors, renaming outcomes in
the database and replacing or removing unknown values. The variables
were either binary or continuous. This code was used on scrambled data
at first so to avoid selection bias as much as possible. When the code
was finished, it was run on the unscrambled data.

@@

@@

As we were running the logistical regression, further were excluded from
the study due to missing data. Among the 8,312 patients screened for
ofi, 7,378 were included in the study. 934 patients were excluded from
the study due to missing values in one or more variables. This left us
with 70 patients in the delayed treatment group and 7,308 patients in
the no delay to treatment group. The flowchart below illustrates the
exclusion and inclusion of patients.

@@

The problem mentioned above with too few patients in the outcome group
made this into a post hoc analysis. As we didn't know the amount of
people who received delayed treatment when we handled the scrambled
data, we ran the code on the real data when we had too many variables
for the dataset. Therefore, we had to change the variables after having
seen the real data, making it a post hoc analysis. This could in theory
open up for possible selection bias, as we might choose variables which
give us results we are looking for. However, all the variables included
in the final analysis were already included in the first one. Variables
were only removed, not added after having seen the real data.

The biases that might have been present in this study would be selection
bias and misclassification bias. Selection bias I've explained above but
misclassification bias is something that we can't control in this study.
As previously mentioned, the possible existence of an ofi is decided
during the MoM conferences, where they then add it to the database.
However, they might not always be right in their judgement and
misclassify from time to time, leading to patients landing in the wrong
group in the analysis. This bias should not be big enough to have a
significant impact on the results, but should be kept in mind.

@@
